---
title: "p8105_hw2_lz2657"
author: Lingyu Zhang
date: Sep 30, 2018
output: github_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

## Problem 1

First, read and clean the data as required.

```{r problem1_read_clean}
NYC_subway_data = read_csv(file = "./data/NYC_Transit_Subway_Entrance_And_Exit_Data.csv") %>%
  janitor::clean_names() %>%
  select(line:entry, vending, ada) %>%
  mutate(entry=as.logical(ifelse(entry == 'YES', TRUE, FALSE)))
```

The dataset now contains variables of line, station name, station latitude / longitude, routes served, entry, vending, entrance type, and ADA compliance. I read the original dataset, cleaned the names, selected the variables I need and convert the entry variable to logical by 'mutate' and 'ifelse' function.

The dimension of the dataset is `r nrow(NYC_subway_data)` x `r ncol(NYC_subway_data)`. The data is not tidy yet.

There are `r nrow(distinct(NYC_subway_data, line, station_name))` distinct stations. 

`r sum(NYC_subway_data$ada==1)` stations are ADA compliant. 

`r round(100*sum(NYC_subway_data$vending=='NO'&NYC_subway_data$entry==TRUE)/sum(NYC_subway_data$vending=='NO'), digits= 2)`% of station entrances / exits without vending allow entrance.